---
title: 'RPII: Legal Identity and Roma Populations in North Macedonia'
author: "Noah Plane"
date: "2024-06-12"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Downloading package facilitating the importing of SPSS files into R

```{r}
if (!requireNamespace("haven", quietly = TRUE)) {
  install.packages("haven", repos = "http://cran.us.r-project.org")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr", repos = "http://cran.us.r-project.org")
}

library(haven)
library(dplyr)
```

# Loading UNICEF MICS North-Macedonia SPSS data form 2018-19 into R all data that is specifically about the Roma population gets the indicator r

```{r}
data_directory <- "~/Desktop/Research Project R Files/RESEARCH-PROJECT-R-FILES/Data/Republic of North Macedonia (Roma Settlements) MICS6 SPSS Datasets"
sav_files <- list.files(data_directory, pattern = "\\.sav$", full.names = TRUE)
datasets <- list()
for (file in sav_files) {
  dataset_name <- paste0(tools::file_path_sans_ext(basename(file)), "r") 
  datasets[[dataset_name]] <- read_sav(file)
}

names(datasets)

data_directory_2 <- "~/Desktop/Research Project R Files/RESEARCH-PROJECT-R-FILES/Data/Republic of North Macedonia MICS6 SPSS Datasets"
sav_files_2 <- list.files(data_directory_2, pattern = "\\.sav$", full.names = TRUE)
for (file in sav_files_2) {
  dataset_name <- tools::file_path_sans_ext(basename(file)) 
  datasets[[dataset_name]] <- read_sav(file)
}
names(datasets)
```

#Checking for if Hl1 needs to be recoded into LN for the hlr dataset

```{r}
names(datasets$hlr)
```

# Recoding Hl1 into LN for the hlr dataset

```{r}
datasets$hlr$LN <- datasets$hlr$HL1
```

# Opening all datasets as a sepreate objects the global environment to check for recoding and cleaning

```{r}
chr <- datasets$chr
hhr <- datasets$hhr
hlr <- datasets$hlr
wmr <- datasets$wmr
bhr <- datasets$bhr
fsr <- datasets$fsr
ch <- datasets$chr
hh <- datasets$hhr
hl <- datasets$hlr
wm <- datasets$wmr
bh <- datasets$bhr
fs <- datasets$fsr
```

# Recoding primary and control variables:

# Recoding indeependant variable

# Creating new variable Index_Birth_Registered_R for the dependant variable that mesures birth registration, combining variable BR1 on birth cerfiticates and BR2 on birth registration into one new binary variable Index_Birth_Registered_R that mesures if the child has a birth certificate or is registered with 1 (BR1=1,2 or BR2=1) beeing yes and 0 (BR1=3 or BR2=2) beeing no with 8 and being missing data (BR1=8 or BR2=8) adding this new variable to the chr dataset next to the original variables BR1 and BR2

```{r}
chr <- chr %>%
  mutate(
    Index_Birth_Registered_R = case_when(
      BR1 %in% c(1, 2) | BR2 == 1 ~ 1,
      BR1 == 3 | BR2 == 2 ~ 0,
      BR1 %in% c(8, NA) | BR2 %in% c(8, NA) ~ NA_real_,
      TRUE ~ NA_real_
    )
  ) %>%
  relocate(Index_Birth_Registered_R, .after = BR2) 

```

# Recoding independent variable

# Creating new variable Index_Disabled_Child_R for the independent variablet hat indicates if a child is disabled (1) or not (0)

# based on a combination of variables related to disabilities (UCF3, UCF4, UCF7, UCF9, UCF11, UCF12, UCF13-UCF18)

# The variable is assigned 1 if any of the disability-related variables indicate disability

# The variable is assigned 0 if none of the disability-related variables indicate disability

# AND at least one variable indicates non-disability

# If all relevant variables are missing, the variable is assigned NA

binary variable Index_Disabled_Child, where 1 is disabled, 0 is not
disabled, and the rest is coded as missing.

```{r}
chr <- chr %>%
  mutate(
    Index_Disabled_Child_R = case_when(
      UCF3 == 1 | UCF4 == 1 | UCF7 %in% c(2, 3, 4) | UCF9 %in% c(2, 3, 4) |
        UCF11 %in% c(2, 3, 4) | UCF12 %in% c(1, 2, 3, 4) |
        UCF13 %in% c(3, 4) | UCF14 %in% c(3, 4) | UCF15 %in% c(3, 4) |
        UCF16 %in% c(3, 4) | UCF17 %in% c(3, 4) | UCF18 %in% c(3, 4) ~ 1,
      (UCF3 == 2 | UCF4 == 2 | UCF7 == 1 | UCF9 == 1 |
        UCF12 == 0 | UCF13 %in% c(1, 2) | UCF14 %in% c(1, 2) |
        UCF15 %in% c(1, 2) | UCF16 %in% c(1, 2) | UCF17 %in% c(1, 2) |
        UCF18 %in% c(1, 2)) &
        !(UCF3 == 1 | UCF4 == 1 | UCF7 %in% c(2, 3, 4) | UCF9 %in% c(2, 3, 4) |
          UCF11 %in% c(2, 3, 4) | UCF12 %in% c(1, 2, 3, 4) |
          UCF13 %in% c(3, 4) | UCF14 %in% c(3, 4) | UCF15 %in% c(3, 4) |
          UCF16 %in% c(3, 4) | UCF17 %in% c(3, 4) | UCF18 %in% c(3, 4)) ~ 0,
      is.na(UCF3) & is.na(UCF4) & is.na(UCF7) & is.na(UCF9) &
        is.na(UCF11) & is.na(UCF12) & is.na(UCF13) & is.na(UCF14) &
        is.na(UCF15) & is.na(UCF16) & is.na(UCF17) & is.na(UCF18) ~ NA_real_,
      TRUE ~ NA_real_
    )
  ) %>%
  relocate(Index_Disabled_Child_R, .after = UCF18)
```

# Recoding independent variable/moderator

# Creating new variable Index_Mother_Disabled_R for the independent variable that indicates if the mother is disabled (1) or not (0)

# based on a combination of variables related to disabilities ((AF2, AF3, AF6, AF8, AF9, AF10, AF11, AF12)

# The variable is assigned 1 if any of the disability-related variables indicate disability

# The variable is assigned 0 if none of the disability-related variables indicate disability

# AND at least one variable indicates non-disability

# If all relevant variables are missing, the variable is assigned NA

```{r}
wmr <- wmr %>%
  mutate(
    Index_Mother_Disabled_R = case_when(
      # Assign 1 (disabled) if any of the variables indicate disability
      AF2 == 1 | AF3 == 1 | AF6 %in% c(2, 3, 4) | AF8 %in% c(2, 3, 4) |
        AF9 %in% c(2, 3, 4) | AF10 %in% c(2, 3, 4) |
        AF11 %in% c(2, 3, 4) | AF12 %in% c(2, 3) ~ 1,
      (AF2 == 2 | AF3 == 2 | AF6 == 1 | AF8 == 1 |
        AF9 == 1 | AF10 == 1 | AF11 == 1 | AF12 == 1) &
        !(AF2 == 1 | AF3 == 1 | AF6 %in% c(2, 3, 4) | AF8 %in% c(2, 3, 4) |
          AF9 %in% c(2, 3, 4) | AF10 %in% c(2, 3, 4) |
          AF11 %in% c(2, 3, 4) | AF12 %in% c(2, 3)) ~ 0,
      is.na(AF2) & is.na(AF3) & is.na(AF6) & is.na(AF8) &
        is.na(AF9) & is.na(AF10) & is.na(AF11) & is.na(AF12) ~ NA_real_,
      TRUE ~ NA_real_
    )
  ) %>%
  relocate(Index_Mother_Disabled_R, .after = AF12)
```

#Sort datasets by key variables HH1, HH2, LN

```{r}
datasets$wmr <- datasets$wmr %>% arrange(HH1, HH2, LN)
datasets$chr <- datasets$chr %>% arrange(HH1, HH2, LN)
```

#Merge data sets by key variables HH1, HH2, LN \# Merge the wmr (women)
and chr (children under 5) datasets \# The relationship between wmr and
chr is one-to-many, where one woman can be the mother/caretaker of
multiple children \# Key variables used for merging: \# - HH1 (cluster
number) \# - HH2 (household number) \# - LN (line number of woman in wmr
dataset) \# - UF4 (line number of mother/caretaker in chr dataset) \#
The chr and wmr datasets where ordered by the key variables before
merging and the variable UF4 was renamed to LN in the wmr dataset to
facilitate the merge

```{r}
wmr <- wmr %>% rename(UF4 = LN)
chr <- chr[order(chr$HH1, chr$HH2, chr$UF4), ]
wmr <- wmr[order(wmr$HH1, wmr$HH2, wmr$UF4), ]
merged_data <- merge(chr, wmr, by = c("HH1", "HH2", "UF4"), all.x = TRUE)
```

# Placing the new index variables at the beginning of the dataset

```{r}
library(dplyr)
merged_data <- merged_data %>                                                             %
  select(HH1, HH2, UF4, Index_Birth_Registered_R, Index_Disabled_Child_R, Index_Mother_Disabled_R)
```

# Save merged dataset for better sharing and in the case of crash

```{r}
write.csv(merged_data, "merged_data.csv", row.names = FALSE)
```

# Start of analysis

# Summary statistics of the merged dataset:

```{r}
summary(merged_data)
```

#Frequency distribution of birth registration:

```{r}
library(ggplot2)

ggplot(merged_data, aes(x = Index_Birth_Registered_R)) +
  geom_bar() +
  labs(title = "Frequency Distribution of Birth Registration",
       x = "Birth Registration Status",
       y = "Count")
```

#Frequency distribution of child disability:

```{r}
ggplot(merged_data, aes(x = Index_Disabled_Child_R)) +
  geom_bar() +
  labs(title = "Frequency Distribution of Child Disability",
       x = "Child Disability Status",
       y = "Count")
```

#Frequency distribution of mother disability:

```{r}
ggplot(merged_data, aes(x = Index_Mother_Disabled_R)) +
  geom_bar() +
  labs(title = "Frequency Distribution of Mother Disability",
       x = "Mother Disability Status",
       y = "Count")
```

#Cross-tabulation of birth registration and child disability:

```{r}
table(merged_data$Index_Birth_Registered_R, merged_data$Index_Disabled_Child_R)
```

#Cross-tabulation of birth registration and mother disability:

```{r}
table(merged_data$Index_Birth_Registered_R, merged_data$Index_Mother_Disabled_R)
```

#Cross-tabulation of child disability and mother disability:

```{r}
table(merged_data$Index_Disabled_Child_R, merged_data$Index_Mother_Disabled_R)
```

#Chi-square test of independence between birth registration and child
disability:

```{r}
chisq.test(merged_data$Index_Birth_Registered_R, merged_data$Index_Disabled_Child_R)
```

#Chi-square test of independence between birth registration and mother
disability:

```{r}
chisq.test(merged_data$Index_Birth_Registered_R, merged_data$Index_Mother_Disabled_R)
```

#Chi-square test of independence between child disability and mother
disability:

```{r}
chisq.test(merged_data$Index_Disabled_Child_R, merged_data$Index_Mother_Disabled_R)
```

#Logistic regression of birth registration on child disability and
mother disability:

```{r}
logit_model1 <- glm(Index_Birth_Registered_R ~ Index_Disabled_Child_R + Index_Mother_Disabled_R,
                   data = merged_data, family = binomial)
summary(logit_model1)
```

#Logistic regression of birth registration on child disability and
mother disability with interaction term:

```{r}
logit_model2 <- glm(Index_Birth_Registered_R ~ Index_Disabled_Child_R * Index_Mother_Disabled_R,
                   data = merged_data, family = binomial)
summary(logit_model2)
```

# Odds ratios and 95% confidence intervals for logit_model1:

```{r}
exp(cbind(OR = coef(logit_model1), confint(logit_model1)))
```

# Odds ratios and 95% confidence intervals for logit_model2:

```{r}
exp(cbind(OR = coef(logit_model2), confint(logit_model2)))
```

# Predicted probabilities for logit_model1:

```{r}
merged_data$Predicted_Probabilities1 <- predict(logit_model1, type = "response")
```

# Predicted probabilities for logit_model2:

```{r}
merged_data$Predicted_Probabilities2 <- predict(logit_model2, type = "response")
```

#Visualizations for Interaction Model

```{r}
ggplot(merged_data, aes(x = Index_Disabled_Child_R, y = Predicted_Probabilities2, color = factor(Index_Mother_Disabled_R))) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(title = "Interaction Model: Predicted Probabilities of Birth Registration",
       x = "Child Disability Status",
       y = "Predicted Probability",
       color = "Mother Disability Status")
```

#Predicted probabilities plot:

```{r}
library(effects)

plot(allEffects(model2), type = "response")
```

#Interaction plot:

```{r}
library(interactions)

interact_plot(model2, pred = Index_Disabled_Child_R, modx = Index_Mother_Disabled_R)
```

# Save merged dataset with predicted probabilities for better sharing and in the case of crash

```{r}
write.csv(merged_data, "merged_data_with_predicted_probabilities.csv", row.names = FALSE)
```

install.packages("dplyr") library(dplyr) hhr \<-
read.csv("path_to_your_file.csv")

# Recoding control variable for religion of the head of the household to determine whether the household follows any religion or follows atheism

# creating new variable HouseHead_Religiosity_R for the control variable

# Based on the values on HC1A, the variable is assigned 0, if the response is =7 (7 represenets no religion)

# Based on HC1A, the variable is assigned 1, if the resposnse is= 1,2,3,4 0r 6 (each resposne corresponds affiliation to a certain religion)

# for any other values, NA is alloted

hh \<- hh %\>% mutate( HouseHead_Religiosity_R = case_when( HC1A %in%
c(1, 2, 3, 4, 6) \~ 1, \# Recode to 1 if HC1A is 1, 2, 3, 4, or 6 HC1A
== 7 \~ 0, \# Recode to 0 if HC1A is 7 TRUE \~ NA_real\_ \# Assign NA
for any other values or missing data ) ) %\>%
relocate(HouseHead_Religiosity_R, .after = HC1B) \# Move the new
variable to be right after HC1B

# creating a new control variable for desire of the last birth desire_for_lastbirth_r

# creating a new variable by using DB1 and DB2

# DB1=CM17 represents if there was a live birth in the last 2 years; if yes, value=1 and if no, value=2

# DB2 represents whether the last preganancy was wanted or not; if yes, value=1 and no, value=2

# the condition for new variable is if CM17=1 and DB2=1, then, desire for birth is yes and is recoded to 1

# condition 2 is, if CM17=1 and DB2=2 then, desire for preganancy is no and is recoeded to 0

# lastly, if the response to CM17=2, then, there was no pregnancy so, the value is recoded to NA.

wm \<- wm %\>% mutate( desire_for_lastbirth_r = case_when( CM17 == 1 &
DB2 == 1 \~ 1, \# If CM17 (DB1) is YES and DB2 is YES, assign 1 CM17 ==
1 & DB2 == 2 \~ 0, \# If CM17 (DB1) is YES and DB2 is NO, assign 0 CM17
== 2 \~ NA_real\_, \# If CM17 (DB1) is NO, assign NA TRUE \~ NA_real\_
\# For any other cases, assign NA ) ) %\>%
relocate(desire_for_lastbirth_r, .after = DB2) \# Relocate the new
variable after DB2

# Creating the dummy variables based on the given subcategories

#Ethnic_Immigration_Dummy: Indicates if the respondent felt discriminated against based on ethnic or immigration origin. 
#Sex_Dummy:Indicates if the respondent felt discriminated against based on sex.
#Sexual_Orientation_Dummy: Indicates if the respondent felt discriminated against based on sexual orientation. #Age_Dummy: Indicates if the respondent felt discriminated against based on age.
#Religion_Belief_Dummy: Indicates if the respondent felt discriminated against based on religion or belief. #Other_Reason_Dummy: Indicates if the respondent felt discriminated against for any other reason. #Assigns 1 if the discrimination is reported (value 1). #Assigns 0 if discrimination is not reported (value 2). #Assigns NA for any other case (e.g., DK or missing values).

wm \<- wm %\>% mutate( Ethnic_Immigration_Dummy = case_when( VT22A == 1
\~ 1, VT22A == 2 \~ 0, TRUE \~ NA_real\_ ), Sex_Dummy = case_when( VT22B
== 1 \~ 1, VT22B == 2 \~ 0, TRUE \~ NA_real\_ ),
Sexual_Orientation_Dummy = case_when( VT22C == 1 \~ 1, VT22C == 2 \~ 0,
TRUE \~ NA_real\_ ), Age_Dummy = case_when( VT22D == 1 \~ 1, VT22D == 2
\~ 0, TRUE \~ NA_real\_ ), Religion_Belief_Dummy = case_when( VT22E == 1
\~ 1, VT22E == 2 \~ 0, TRUE \~ NA_real\_ ), Other_Reason_Dummy =
case_when( VT22X == 1 \~ 1, VT22X == 2 \~ 0, TRUE \~ NA_real\_ ) ) %\>%
relocate(Ethnic_Immigration_Dummy, .after = VT22X) %\>%
relocate(Sex_Dummy, .after = Ethnic_Immigration_Dummy) %\>%
relocate(Sexual_Orientation_Dummy, .after = Sex_Dummy) %\>%
relocate(Age_Dummy, .after = Sexual_Orientation_Dummy) %\>%
relocate(Religion_Belief_Dummy, .after = Age_Dummy) %\>%
relocate(Other_Reason_Dummy, .after = Religion_Belief_Dummy)

# Creating a new variable for education level among mothers: Create a new variable level_of_education_mother_r.

# If CM1 is 1 (mother) and WB6A is 3, 4, 5, or 6 (higher education levels), set to 1 (educated): Set to 1 if the woman is a mother and has a higher level of education.

# If CM1 is 1 (mother) and WB6A is 1, 2, or 000 (lower education levels), set to 0 (not educated): Set to 0 if the woman is a mother and has a lower level of education.

# If CM1 is 2 (not a mother), set to NA (not applicable): Set to NA if the woman is not a mother.

# For any other case, set to NA: Default to NA for any other cases.

# Relocate the new variable after the WB6A column: Move the new variable to be after the WB6A column.

wm \<- wm %\>% mutate( level_of_education_mother_r = case_when( CM1 == 1
& WB6A %in% c(3, 4, 5, 6) \~ 1, \# If CM1=1 and WB6A is 3, 4, 5, 6, then
educated (1) CM1 == 1 & WB6A %in% c(1, 2, 000) \~ 0, \# If CM1=1 and
WB6A is 1, 2, or 000, then not educated (0) CM1 == 2 \~ NA_real\_, \# If
CM1=2, then NA (not a mother) TRUE \~ NA_real\_ \# Any other case,
assign NA ) ) %\>% relocate(level_of_education_mother_r, .after = WB6A)

#Create a new variable called Mother_Health_Insurance. \
# If a woman is a mother (CM1 == 1) and has health insurance (WB18 == 1), set Mother_Health_Insurance to 1 (insured). 
# If a woman is a mother (CM1== 1) and does not have health insurance (WB18 == 2), set Mother_Health_Insurance to 0 (not insured). 
# If a woman is not a mother (CM1 == 2), set Mother_Health_Insurance to NA (not applicable).
# For any other case, set Mother_Health_Insurance to NA. 
# Relocating the Mother_Health_Insurance variable to be immediately after the WB18 column.

wm \<- wm %\>% mutate( \# Creating a new variable for health insurance
status among mothers Mother_Health_Insurance = case_when( \# If CM1 is 1
(mother) and WB18 is 1 (has health insurance), set to 1 (insured) CM1 ==
1 & WB18 == 1 \~ 1,\
\# If CM1 is 1 (mother) and WB18 is 2 (no health insurance), set to 0
(not insured) CM1 == 1 & WB18 == 2 \~ 0,\
\# If CM1 is 2 (not a mother), set to NA (not applicable) CM1 == 2 \~
NA_real\_,\
\# For any other case, set to NA TRUE \~ NA_real\_\
) ) %\>% \# Relocate the new variable after the WB18 column
relocate(Mother_Health_Insurance, .after = WB18)

hh <- hh %>%
  rename(ST3_1 = `ST3$1`)
  
  hh <- hh %>%
  rename(ST3_2 = `ST3$2`)
  hh <- hh %>%
  rename(ST3_3 = `ST3$3`)
  hh <- hh %>%
  rename(ST3_4 = `ST3$4`)
  hh <- hh %>%
  rename(ST3_5 = `ST3$5`)
  hh <- hh %>%
  rename(ST3_6 = `ST3$6`)

# Create the new variable for family economic status based on social assistance responses
hh <- hh %>%
  mutate(
    # Calculate the number of 'yes' responses for each family
    num_assistance_yes = (ST3_1 == 1) + (ST3_2 == 1) + (ST3_3 == 1) + 
                         (ST3_4 == 1) + (ST3_5 == 1) + (ST3_6 == 1),
    
    # Classify the families based on the number of 'yes' responses
    family_economic_status = case_when(
      # Poor: Yes to 3 or more programs
      num_assistance_yes >= 3 ~ 0,
      
      # Average: Yes to 1 or 2 programs
      num_assistance_yes >= 1 & num_assistance_yes <= 2 ~ 1,
      
      # Rich: No to all programs
      num_assistance_yes == 0 ~ 2,
      
      # Any other case should be treated as NA (not applicable here, but added for completeness)
      TRUE ~ NA_real_
    )
  ) %>%
  # Remove the temporary num_assistance_yes column
  select(-num_assistance_yes) %>%
  # Relocate the new variable after the last ST3 column
  relocate(family_economic_status, .after = ST3_6)
  
  