---
title: 'RPII: Legal Identity and Roma Populations in North Macedonia'
author: "Noah Plane"
date: "2024-06-12"
output: html_document
---

# Downloading package facilitating the importing of SPSS files into R

```{r}
if (!requireNamespace("haven", quietly = TRUE)) {
  install.packages("haven", repos = "http://cran.us.r-project.org")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr", repos = "http://cran.us.r-project.org")
}

library(haven)
library(dplyr)
```


# Loading UNICEF MICS North-Macedonia SPSS data form 2018-19 into R all data that is specifically about the Roma population gets the indicator r

```{r}
data_directory <- "~/Desktop/Research Project R Files/RESEARCH-PROJECT-R-FILES/Data/Republic of North Macedonia (Roma Settlements) MICS6 SPSS Datasets"
sav_files <- list.files(data_directory, pattern = "\\.sav$", full.names = TRUE)
datasets <- list()
for (file in sav_files) {
  dataset_name <- paste0(tools::file_path_sans_ext(basename(file)), "r") 
  datasets[[dataset_name]] <- read_sav(file)
}

names(datasets)

data_directory_2 <- "~/Desktop/Research Project R Files/RESEARCH-PROJECT-R-FILES/Data/Republic of North Macedonia MICS6 SPSS Datasets"
sav_files_2 <- list.files(data_directory_2, pattern = "\\.sav$", full.names = TRUE)
for (file in sav_files_2) {
  dataset_name <- tools::file_path_sans_ext(basename(file)) 
  datasets[[dataset_name]] <- read_sav(file)
}
names(datasets)
```




#Checking for if Hl1 needs to be recoded into LN for the hlr dataset

```{r}
names(datasets$hlr)
```

# Recoding Hl1 into LN for the hlr dataset
```{r}
datasets$hlr$LN <- datasets$hlr$HL1
```

# Opening all datasets as a sepreate objects the global environment to check for recoding and cleaning
```{r}
chr <- datasets$chr
hhr <- datasets$hhr
hlr <- datasets$hlr
wmr <- datasets$wmr
bhr <- datasets$bhr
fsr <- datasets$fsr
ch <- datasets$chr
hh <- datasets$hhr
hl <- datasets$hlr
wm <- datasets$wmr
bh <- datasets$bhr
fs <- datasets$fsr
```

# Recoding primary and control variables:

# Recoding indeependant variable
# Creating new variable Index_Birth_Registered_R for the dependant variable that mesures birth registration, combining variable BR1 on birth cerfiticates and BR2 on birth registration into one new binary variable Index_Birth_Registered_R that mesures if the child has a birth certificate or is registered with 1 (BR1=1,2 or BR2=1) beeing yes and 0 (BR1=3 or BR2=2) beeing no with 8 and being missing data (BR1=8 or BR2=8) adding this new variable to the chr dataset next to the original variables BR1 and BR2
```{r}
chr <- chr %>%
  mutate(
    Index_Birth_Registered_R = case_when(
      BR1 %in% c(1, 2) | BR2 == 1 ~ 1,
      BR1 == 3 | BR2 == 2 ~ 0,
      BR1 %in% c(8, NA) | BR2 %in% c(8, NA) ~ NA_real_,
      TRUE ~ NA_real_
    )
  ) %>%
  relocate(Index_Birth_Registered_R, .after = BR2) 

```

# Recoding independent variable 
# Creating new variable Index_Disabled_Child_R for the independent variablet hat indicates if a child is disabled (1) or not (0) 
# based on a combination of variables related to disabilities (UCF3, UCF4, UCF7, UCF9, UCF11, UCF12, UCF13-UCF18)
# The variable is assigned 1 if any of the disability-related variables indicate disability
# The variable is assigned 0 if none of the disability-related variables indicate disability
# AND at least one variable indicates non-disability
# If all relevant variables are missing, the variable is assigned NA

binary variable Index_Disabled_Child, where 1 is disabled, 0 is not disabled, and the rest is coded as missing.
```{r}
chr <- chr %>%
  mutate(
    Index_Disabled_Child_R = case_when(
      UCF3 == 1 | UCF4 == 1 | UCF7 %in% c(2, 3, 4) | UCF9 %in% c(2, 3, 4) |
        UCF11 %in% c(2, 3, 4) | UCF12 %in% c(1, 2, 3, 4) |
        UCF13 %in% c(3, 4) | UCF14 %in% c(3, 4) | UCF15 %in% c(3, 4) |
        UCF16 %in% c(3, 4) | UCF17 %in% c(3, 4) | UCF18 %in% c(3, 4) ~ 1,
      (UCF3 == 2 | UCF4 == 2 | UCF7 == 1 | UCF9 == 1 |
        UCF12 == 0 | UCF13 %in% c(1, 2) | UCF14 %in% c(1, 2) |
        UCF15 %in% c(1, 2) | UCF16 %in% c(1, 2) | UCF17 %in% c(1, 2) |
        UCF18 %in% c(1, 2)) &
        !(UCF3 == 1 | UCF4 == 1 | UCF7 %in% c(2, 3, 4) | UCF9 %in% c(2, 3, 4) |
          UCF11 %in% c(2, 3, 4) | UCF12 %in% c(1, 2, 3, 4) |
          UCF13 %in% c(3, 4) | UCF14 %in% c(3, 4) | UCF15 %in% c(3, 4) |
          UCF16 %in% c(3, 4) | UCF17 %in% c(3, 4) | UCF18 %in% c(3, 4)) ~ 0,
      is.na(UCF3) & is.na(UCF4) & is.na(UCF7) & is.na(UCF9) &
        is.na(UCF11) & is.na(UCF12) & is.na(UCF13) & is.na(UCF14) &
        is.na(UCF15) & is.na(UCF16) & is.na(UCF17) & is.na(UCF18) ~ NA_real_,
      TRUE ~ NA_real_
    )
  ) %>%
  relocate(Index_Disabled_Child_R, .after = UCF18)
```

# Recoding independent variable/moderator 
# Creating new variable Index_Mother_Disabled_R for the independent variable that indicates if the mother is disabled (1) or not (0)
# based on a combination of variables related to disabilities ((AF2, AF3, AF6, AF8, AF9, AF10, AF11, AF12)
# The variable is assigned 1 if any of the disability-related variables indicate disability
# The variable is assigned 0 if none of the disability-related variables indicate disability
# AND at least one variable indicates non-disability
# If all relevant variables are missing, the variable is assigned NA
```{r}
wmr <- wmr %>%
  mutate(
    Index_Mother_Disabled_R = case_when(
      # Assign 1 (disabled) if any of the variables indicate disability
      AF2 == 1 | AF3 == 1 | AF6 %in% c(2, 3, 4) | AF8 %in% c(2, 3, 4) |
        AF9 %in% c(2, 3, 4) | AF10 %in% c(2, 3, 4) |
        AF11 %in% c(2, 3, 4) | AF12 %in% c(2, 3) ~ 1,
      (AF2 == 2 | AF3 == 2 | AF6 == 1 | AF8 == 1 |
        AF9 == 1 | AF10 == 1 | AF11 == 1 | AF12 == 1) &
        !(AF2 == 1 | AF3 == 1 | AF6 %in% c(2, 3, 4) | AF8 %in% c(2, 3, 4) |
          AF9 %in% c(2, 3, 4) | AF10 %in% c(2, 3, 4) |
          AF11 %in% c(2, 3, 4) | AF12 %in% c(2, 3)) ~ 0,
      is.na(AF2) & is.na(AF3) & is.na(AF6) & is.na(AF8) &
        is.na(AF9) & is.na(AF10) & is.na(AF11) & is.na(AF12) ~ NA_real_,
      TRUE ~ NA_real_
    )
  ) %>%
  relocate(Index_Mother_Disabled_R, .after = AF12)
```



#Sort datasets by key variables HH1, HH2, LN

```{r}
datasets$wmr <- datasets$wmr %>% arrange(HH1, HH2, LN)
datasets$chr <- datasets$chr %>% arrange(HH1, HH2, LN)
```

#Merge data sets by key variables HH1, HH2, LN
# Merge the wmr (women) and chr (children under 5) datasets
# The relationship between wmr and chr is one-to-many, where one woman can be the mother/caretaker of multiple children
# Key variables used for merging:
#   - HH1 (cluster number)
#   - HH2 (household number)
#   - LN (line number of woman in wmr dataset)
#   - UF4 (line number of mother/caretaker in chr dataset)
# The chr and wmr datasets where ordered by the key variables before merging and the variable UF4 was renamed to LN in the wmr dataset to facilitate the merge 
```{r}
wmr <- wmr %>% rename(UF4 = LN)
chr <- chr[order(chr$HH1, chr$HH2, chr$UF4), ]
wmr <- wmr[order(wmr$HH1, wmr$HH2, wmr$UF4), ]
merged_data <- merge(chr, wmr, by = c("HH1", "HH2", "UF4"), all.x = TRUE)
```



# Placing the new index variables at the beginning of the dataset 
```{r}
library(dplyr)
merged_data <- merged_data %>                                                             %
  select(HH1, HH2, UF4, Index_Birth_Registered_R, Index_Disabled_Child_R, Index_Mother_Disabled_R)
```


# Save merged dataset for better sharing and in the case of crash
```{r}
write.csv(merged_data, "merged_data.csv", row.names = FALSE)
```


# Start of analysis
# Summary statistics of the merged dataset:
```{r}
summary(merged_data)
```

#Frequency distribution of birth registration:
```{r}
library(ggplot2)

ggplot(merged_data, aes(x = Index_Birth_Registered_R)) +
  geom_bar() +
  labs(title = "Frequency Distribution of Birth Registration",
       x = "Birth Registration Status",
       y = "Count")
```


#Frequency distribution of child disability:
```{r}
ggplot(merged_data, aes(x = Index_Disabled_Child_R)) +
  geom_bar() +
  labs(title = "Frequency Distribution of Child Disability",
       x = "Child Disability Status",
       y = "Count")
```

#Frequency distribution of mother disability:
```{r}
ggplot(merged_data, aes(x = Index_Mother_Disabled_R)) +
  geom_bar() +
  labs(title = "Frequency Distribution of Mother Disability",
       x = "Mother Disability Status",
       y = "Count")
```

#Cross-tabulation of birth registration and child disability:
```{r}
table(merged_data$Index_Birth_Registered_R, merged_data$Index_Disabled_Child_R)
```

#Cross-tabulation of birth registration and mother disability:
```{r}
table(merged_data$Index_Birth_Registered_R, merged_data$Index_Mother_Disabled_R)
```

#Cross-tabulation of child disability and mother disability:
```{r}
table(merged_data$Index_Disabled_Child_R, merged_data$Index_Mother_Disabled_R)
```

#Chi-square test of independence between birth registration and child disability:
```{r}
chisq.test(merged_data$Index_Birth_Registered_R, merged_data$Index_Disabled_Child_R)
```

#Chi-square test of independence between birth registration and mother disability:
```{r}
chisq.test(merged_data$Index_Birth_Registered_R, merged_data$Index_Mother_Disabled_R)
```

#Chi-square test of independence between child disability and mother disability:
```{r}
chisq.test(merged_data$Index_Disabled_Child_R, merged_data$Index_Mother_Disabled_R)
```

#Logistic regression of birth registration on child disability and mother disability:
```{r}
logit_model1 <- glm(Index_Birth_Registered_R ~ Index_Disabled_Child_R + Index_Mother_Disabled_R,
                   data = merged_data, family = binomial)
summary(logit_model1)
```

#Logistic regression of birth registration on child disability and mother disability with interaction term:
```{r}
logit_model2 <- glm(Index_Birth_Registered_R ~ Index_Disabled_Child_R * Index_Mother_Disabled_R,
                   data = merged_data, family = binomial)
summary(logit_model2)
```


# Odds ratios and 95% confidence intervals for logit_model1:
```{r}
exp(cbind(OR = coef(logit_model1), confint(logit_model1)))
```

# Odds ratios and 95% confidence intervals for logit_model2:
```{r}
exp(cbind(OR = coef(logit_model2), confint(logit_model2)))
```

# Predicted probabilities for logit_model1:
```{r}
merged_data$Predicted_Probabilities1 <- predict(logit_model1, type = "response")
```

# Predicted probabilities for logit_model2:
```{r}
merged_data$Predicted_Probabilities2 <- predict(logit_model2, type = "response")
```

#Visualizations for Interaction Model
```{r}
ggplot(merged_data, aes(x = Index_Disabled_Child_R, y = Predicted_Probabilities2, color = factor(Index_Mother_Disabled_R))) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(title = "Interaction Model: Predicted Probabilities of Birth Registration",
       x = "Child Disability Status",
       y = "Predicted Probability",
       color = "Mother Disability Status")
```

#Predicted probabilities plot:
```{r}
library(effects)

plot(allEffects(model2), type = "response")
```

#Interaction plot:
```{r}
library(interactions)

interact_plot(model2, pred = Index_Disabled_Child_R, modx = Index_Mother_Disabled_R)
```

# Save merged dataset with predicted probabilities for better sharing and in the case of crash
```{r}
write.csv(merged_data, "merged_data_with_predicted_probabilities.csv", row.names = FALSE)
```





































